---
title: "Data Analysis for Fibrosis paper - Loading and initial processing of stroma"
author: "James Reineke and Matt Cannon"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
format:
    html:
        toc: true
        toc_float: true
        toc_depth: 5
        number_sections: false
        code-fold: true
        code-link: true
        df-print: kable
        embed-resources: true
execute:
    cache: true
knitr:
    opts_chunk:
        cache.lazy: false
---

# Sample summary

- Our data
    - stuff here

# Load packages
```{r lib}
library(cowplot)
library(cluster)
library(tidyverse)
library(ggrepel)
library(patchwork)
library(reticulate)
library(Seurat)
library(sctransform)
library(rrrSingleCellUtils)
library(msigdbr)
library(parallel)

# Set random generator seed to facilitate reproducibility
set.seed(888)
```

## Make up directory structure
```{bash mkdirs, eval=TRUE}
for directoryName in \
  output \
  output/figures \
  output/figures/spatial \
  output/rdata \
  output/de
do
  if [ ! -d ${directoryName} ]
  then
    mkdir -p ${directoryName}
  fi
done
```

## Load functions

{{< include child_qmds/helper_functions.qmd >}}

## Read in all data and process it for downstream analysis
List of seurat object saved to `output/rdata/sobj_list.qs`

{{< include child_qmds/Load-Cluster.qmd >}}

## Analyze C57BL/6 and F420 data

{{< include child_qmds/b6_f420_analysis.qmd >}}

## Analyze BALBC and K7M2 data

{{< include child_qmds/balbc_k7m2_analysis.qmd >}}

## Analyze macrophage populations within the tumor samples
This includes both metastatic and primary tumor samples
Merged object saved to `output/rdata/all_sobj.qs`
The subset macrophages from lung mets and primary tumors are saved to `output/rdata/immune_recurl.qs`

{{< include child_qmds/Epithelial-Macs.qmd >}}

## Run GSEA on the macrophage populations

{{< include child_qmds/gsea_macs.qmd >}}

## Load in single Visium lung met sample

{{< include child_qmds/spatial_load.qmd >}}

## Generate a reference of F420 tumor cells from lung metastasis

{{< include child_qmds/F420_ref.qmd >}}

## Run cell deconvolution on spatial data with spacexr

{{< include child_qmds/spatial_spacexr.qmd >}}

## Look at spatial correlation of cell types

{{< include child_qmds/spatial_co_occur.qmd >}}

```{r sessionInfo}
sessionInfo()
```
