--
title: "Data Analysis for Fibrosis paper"
author: "James Reineke and Matt Cannon"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
    html_document:
        toc: true
        toc_float: true
        toc_depth: 5
        number_sections: false
        code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    tidy = TRUE,
    echo = TRUE,
    cache = TRUE,
    collapse = TRUE,
    tidy.opts = list(width.cutoff = 95),
    cache.lazy = FALSE
)
```

# Load packages
```{r lib}
library(cowplot)
library(cluster)
library(tidyverse)
library(ggrepel)
library(patchwork)
library(reticulate)
library(Seurat)
library(sctransform)
library(rrrSingleCellUtils)
library(msigdbr)
# Set random generator seed to facilitate reproducibility
set.seed(888)
```

## Make up directory structure
```{bash mkdirs, eval=TRUE}
for directoryName in \
  output \
  output/figures \
  output/rdata \
  output/de
do
  if [ ! -d ${directoryName} ]
  then
    mkdir -p ${directoryName}
  fi
done
```
 
# Load raw data
```{r load_raw, eval=TRUE}
sample_list <- read_tsv("misc/sample_cutoffs.txt", show_col_types = FALSE)

sobj_list <- list()

end_path <- "/filtered_feature_bc_matrix"
if (grepl("r1pl", Sys.info()[["nodename"]])) {
    data_path <- "/home/gdrobertslab/lab/Counts/"
    path_use <- "cluster_path"
} else {
    data_path <- "/Applications/scRNA_seq raw files/"
    path_use <- "James_path"
}

for (i in seq_len(nrow(sample_list))) {
    obj_name <- sample_list$obj_name[i]

    sobj_list[[obj_name]] <-
        tenx_load_qc(paste0(data_path,
                            sample_list[[path_use]][i],
                            end_path),
                     min_cells = 3,
                     min_features = 200,
                     violin_plot = FALSE)

    # Add metadata to dataset
    for (colname in colnames(sample_list)) {
        sobj_list[[obj_name]][[colname]] <- sample_list[[colname]][i]
    }

    # Add sample name to dataset
    sobj_list[[obj_name]]$sample <- obj_name

    cutoff_table <-
        tribble(~"feature",   ~"min_val",                    ~"max_val",
                "nCount_RNA", sample_list$ncount_rna_min[i], sample_list$ncount_rna_max[i],
                "percent.mt", 0,                             sample_list$mt_max[i])

    plotted <-
        feature_hist(sobj_list[[obj_name]],
                     features = c("nCount_RNA", "percent.mt"),
                     cutoff_table = cutoff_table)

    ggsave(paste0("output/figures/feature_hist_",
                  obj_name,
                  ".png"),
           width = 10,
           height = 10,
           plot = plotted)

    sobj_list[[obj_name]] <-
        sobj_list[[obj_name]] %>%
        subset(nCount_RNA   >= sample_list$ncount_rna_min[i] &
               nCount_RNA   <= sample_list$ncount_rna_max[i] &
               percent.mt   <= sample_list$mt_max[i]) %>%
        process_seurat()

}
save(sobj_list,
     file = "output/rdata/sobj_list.rds")
```


## Combine B6 and F420
```{r b6_f420, dependson='load_raw', eval=TRUE}
###scTransform merged data sets to include all genes in scaled data 
B6.f420.combined <-
    merge(sobj_list[["C57BL6"]],
          y = sobj_list[["F420"]],
          add.cell.ids = c("C57BL6", "F420")) %>%
    SCTransform(vars.to.regress = "percent.mt",
                verbose = FALSE,
                return.only.var.genes = FALSE) %>%
    process_seurat(resolution = 0.2,
                   run_umap_dims = 1:30)

DimPlot(B6.f420.combined,
        label = TRUE,
        split.by = "sample",
        pt.size = 1,
        cols = scales::alpha(plot_cols, 0.6)) +
    coord_fixed() +
    theme(axis.text = element_text(size = 5),
          axis.title = element_text(size = 8),
          plot.title = element_text(size = 10,
                                    face = "bold",
                                    hjust = 0.5),
          strip.text = element_text(size = 8, face = "bold"),
          plot.subtitle = element_text(size = 8, hjust = 0.5),
          legend.text = element_text(size = 5),
          legend.title = element_text(size = 6, face = "bold"),
          strip.background = element_blank())

ggsave("output/figures/combined_B6_F420.pdf",
       width = 4,
       height = 4)
```

## Combine BALBC and K7M2
```{r balbc_k7m2, dependson='load_raw', eval=TRUE}
###merge control and met datasets 
balbc.k7m2.combined <-
    merge(sobj_list[["BALBC"]],
          y = sobj_list[["K7M2"]],
          add.cell.ids = c("BALBC", "K7M2")) %>%
    SCTransform(vars.to.regress = "percent.mt",
                verbose = FALSE,
                return.only.var.genes = FALSE) %>%
    process_seurat(resolution = 0.2,
                   run_umap_dims = 1:30)

###Remove  dead cells cluster 9  noted by high expression of mitochondrial as well as small clusters 18-20

#balbc.k7m2.combined.subset<-subset(balbc.k7m2.combined, idents = c("9", "18", "19", "20"), invert=T)
balbc.k7m2.combined.subset<-RenameIdents(balbc.k7m2.combined.subset, `0`="Alveolar macrophage", `1`="Macrophage",`2`="T cell",`3`="Endothelial cell", `4`="K7M2", `5`="Granulocyte", `6`="Dendritic cell", `7`="B cell", `8`="Distal airway cell", `10`="Fibroblast", `11`="Endothelial cell", `12`="Macrophage",`13`="T cell",  `14`=" Proximal airway cell", `15`="Alveolar macrophage", `16`="Endothelial cell", `17`="Adipocyte")
###add cluster identity to metadata
balbc.k7m2.combined.subset$celltype.sample <- paste(Idents(balbc.k7m2.combined.subset), balbc.k7m2.combined.subset$sample, sep = "_") 
balbc.k7m2.combined.subset$celltype <- Idents(balbc.k7m2.combined.subset)

```

### Run DE on combined data set
```{r de_b6_F420, dependson='b6_f420', eval=TRUE}
initial.markers <-
    FindAllMarkers(B6.f420.combined,
                   only.pos = TRUE,
                   min.pct = 0.25,
                   logfc.threshold = 0.5)
```

### Remove cycling and dead cells 
clusters 3 and 13 noted by high expression of mitochondrial and cell cycling genes respectively

We should come up with automated ways to define these - MVC
```{r remove_cycling, dependson='de_b6_F420', eval=TRUE}
B6.f420.combined.subset <-
    subset(B6.f420.combined,
           idents = c("3", "13"),
           invert = TRUE)
```

### Cluster assignments
Based on canonical markers and marker genes.
```{r cluster_assignments, dependson='remove_cycling', eval=TRUE}
B6.f420.combined.subset <-
    RenameIdents(B6.f420.combined.subset,
                 `0` = "Macrophage",
                 `1` = "Alveolar macrophage",
                 `2` = "Endothelial cell",
                 `4` = "T cell",
                 `5` = "Dendritic cell",
                 `6` = "Proximal airway cell",
                 `7` = "Distal airway cell",
                 `8` = "Fibroblast",
                 `9` = "B cell",
                 `10` = "Endothelial cell",
                 `11` = "Dendritic cell",
                 `12` = "Granulocyte",
                 `14` = "F420",
                 `15` = "Erythrocyte",
                 `16` = "Smooth muscle cell",
                 `17` = "Adipocyte",
                 `18` = "Mesothelial cell")

###add cluster identity to metadata
B6.f420.combined.subset$celltype.sample <-
    paste(Idents(B6.f420.combined.subset),
          B6.f420.combined.subset$sample,
          sep = "_")

B6.f420.combined.subset$celltype <- Idents(B6.f420.combined.subset)
```

### Addmodules GSEA with msigdbr
```{r add_modules, dependson='cluster_assignments', eval=TRUE}
###markers for plotting
B6.F420.combined.plotmarkers <-
    c("Itgam", "Fcgr1", "Cd68", "Adgre1", "Ccr2", "Cx3cr1", "Ly6c2", "Itgax",
      "Mertk", "Marco", "Siglecf", "Egr2", "Bhlhe41", "Fabp1", "Olr1", "Alox5",
      "Vwf", "Tek", "Acer2", "Cavin3", "Pecam1", "Emcn", "Lyve1", "Ptprb",
      "Plvap", "Cd3d", "Cd4", "Cd8b1", "Tcf7", "Il7r", "Lef1", "Dusp10", "Itk",
      "Cd207", "Cd209a", "Ccr7", "Ccl17", "Ccl22", "Wdfy4", "Clec4b1",
      "H2-DMb1", "Etv3", "Itgae", "Xcr1", "Epcam", "Scgb1a1", "Tuba1a", "Foxj1",
      "Sftpd", "Lamp3", "Etv5", "Ager", "Aqp5", "Col1a2", "Col3a1", "Col1a1",
      "Loxl1", "Pdgfra", "Pdgfrb", "Fn1", "Ltbp4", "Bgn", "Cd19", "Cd79a",
      "Cd79b", "Iglc2", "Ighm", "Pax5", "Cd33", "Csf3r", "S100a9", "Cxcr2",
      "Mxd1", "Ptafr", "Steap1", "Crabp2", "Wisp1", "Dclk1", "Twist1", "Il1rl1",
      "Ctgf", "Hbb-bs", "Hba-a1", "Myh11", "Cox4i2", "Pde5a", "Cfd", "Car3",
      "Adipoq", "Msln")

all_gene_sets <- msigdbr(species = "Mus musculus")

pathways <-
    list("HALLMARK_INFLAMMATORY_RESPONSE" = "Inflammation_features",
         "WP_LUNG_FIBROSIS" = "Lung_fibrosis_features",
         "HALLMARK_TNFA_SIGNALING_VIA_NFKB" = "TNFA_features",
         "REACTOME_DEGRADATION_OF_THE_EXTRACELLULAR_MATRIX" = "ECM_degrad_features",
         "REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION" = "ECM_org_features",
         "GOBP_RESPONSE_TO_WOUNDING" = "Wound_response_features")

for (pathway in pathways) {
    gene_list <-
        list(filter(all_gene_sets,
               gs_name == pathway)$gene_symbol)
    B6.f420.combined.subset <-
        AddModuleScore(B6.f420.combined.subset,
                       features = gene_list,
                       name = pathways[[pathway]])
}

```
###subset epithelial cells and re-normalize
b6.f420.combined.AEC=subset(B6.f420.combined.subset, idents = "Distal airway cell")
b6.f420.combined.AEC<- SCTransform(b6.f420.combined.AEC, vars.to.regress = "percent.mt", verbose = FALSE, return.only.var.genes = FALSE)
###dim reduction
b6.f420.combined.AEC<- RunPCA(b6.f420.combined.AEC, verbose = FALSE)
b6.f420.combined.AEC<- RunUMAP(b6.f420.combined.AEC, dims = 1:30, verbose = FALSE)
###cluster
b6.f420.combined.AEC<- FindNeighbors(b6.f420.combined.AEC, dims = 1:30, verbose = FALSE)
b6.f420.combined.AEC<- FindClusters(b6.f420.combined.AEC,resolution=0.3, verbose = FALSE)
###remove contaminating non-epithelial cells-immune cells
b6.f420.combined.AEC=subset(b6.f420.combined.AEC, idents = c("1","2", "3", "6"))
###re-cluster with pure epithelial cell populations 
b6.f420.combined.AEC_1<- SCTransform(b6.f420.combined.AEC, vars.to.regress = "percent.mt", verbose = FALSE, return.only.var.genes = FALSE)
b6.f420.combined.AEC_1<- RunPCA(b6.f420.combined.AEC_1, verbose = FALSE)
b6.f420.combined.AEC_1<- RunUMAP(b6.f420.combined.AEC_1, dims = 1:30, verbose = FALSE)
b6.f420.combined.AEC_1<- FindNeighbors(b6.f420.combined.AEC_1, dims = 1:30, verbose = FALSE)
b6.f420.combined.AEC_1<- FindClusters(b6.f420.combined.AEC_1,resolution=0.6, verbose = FALSE)
initial.epimarkers <- FindAllMarkers(b6.f420.combined.AEC_1, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.5)
###remove contaminating proximal airway cells
b6.f420.combined.AEC_1=subset(b6.f420.combined.AEC_1, idents = "5", invert=T)
###cluster identification based on published data sets-PMIDs 32750316,32661339,32678092 
b6.f420.combined.AEC_1<-RenameIdents(b6.f420.combined.AEC_1, `0`="pAEC2", `1`="DATP", `2`="AEC2", `3`="AEC1", `4`="pAEC2", `6`="AEC1", `7`="cAEC2")
b6.f420.combined.AEC_1$celltype.sample <- paste(Idents(b6.f420.combined.AEC_1), b6.f420.combined.AEC_1$sample, sep = "_") 
b6.f420.combined.AEC_1$celltype <- Idents(b6.f420.combined.AEC_1)
### add module for DATP
p53_list<-filter(all_gene_sets, gs_name=="HALLMARK_P53_PATHWAY")
p53_list<-list(c(p53_list$gene_symbol))
b6.f420.combined.AEC_1<-AddModuleScore(b6.f420.combined.AEC_1, features =p53_list, name = 'p53_features')
senescence_list<-filter(all_gene_sets, gs_name=="FRIDMAN_SENESCENCE_UP")
senescence_list<-list(c(senescence_list$gene_symbol))
b6.f420.combined.AEC_1<-AddModuleScore(b6.f420.combined.AEC_1, features =senescence_list, name = 'senescence_features')
Distalairway.markers<-c("Lcn2","Lrg1","Cxcl17","Glrx","Krt8","Sfn","Krt19","Cldn4","Trp53","Ctgf","Sftpa1","Sftpb","Sftpc","Lpcat1","Etv5", "Abca3", "Pdpn", "Aqp5", "Hopx", "Cav1","Ager", "Akap5", "Clic5", "Top2a", "Mki67", "Cdk1")
###scTransform macs 
f420.macs<- SCTransform(f420.macs, vars.to.regress = "percent.mt", verbose = FALSE, return.only.var.genes = FALSE)
###dim reduction
f420.macs<- RunPCA(f420.macs, verbose = FALSE)
f420.macs<- RunUMAP(f420.macs, dims = 1:30, verbose = FALSE)
###cluster 
f420.macs<- FindNeighbors(f420.macs, dims = 1:30, verbose = FALSE)
f420.macs<- FindClusters(f420.macs,resolution=0.4, verbose = FALSE)
mac.markers<-c("Itgam","Itgax", "Cd14", "Fcgr3", "Fcgr1", "Cd68", "Adgre1", "Ccr2", "Cx3cr1", "Ly6c1", "Ly6c2", "H2-Ab1", "H2-DMb1")
###Create and merge BALBC and k7M2 seurat objects followed by scTransform


###merge control and met datasets 
balbc.k7m2.combined=merge(BALBC, y=K7M2, add.cell.ids = c("BALBC","K7M2"))
###scTransform merged data sets to include all genes in scaled data 
balbc.k7m2.combined<- SCTransform(balbc.k7m2.combined, vars.to.regress = "percent.mt", verbose = FALSE, return.only.var.genes = FALSE)
###dim reduction
balbc.k7m2.combined <- RunPCA(balbc.k7m2.combined, verbose = FALSE)
balbc.k7m2.combined <- RunUMAP(balbc.k7m2.combined, dims = 1:30, verbose = FALSE)
###cluster low resolution "birds eye view" of lung microenvironment
balbc.k7m2.combined<- FindNeighbors(balbc.k7m2.combined, dims = 1:30, verbose = FALSE)
balbc.k7m2.combined<- FindClusters(balbc.k7m2.combined,resolution=0.2, verbose = FALSE)
###Remove  dead cells cluster 9  noted by high expression of mitochondrial as well as small clusters 18-20
balbc.k7m2.combined.subset<-subset(balbc.k7m2.combined, idents = c("9", "18", "19", "20"), invert=T)
balbc.k7m2.combined.subset<-RenameIdents(balbc.k7m2.combined.subset, `0`="Alveolar macrophage", `1`="Macrophage",`2`="T cell",`3`="Endothelial cell", `4`="K7M2", `5`="Granulocyte", `6`="Dendritic cell", `7`="B cell", `8`="Distal airway cell", `10`="Fibroblast", `11`="Endothelial cell", `12`="Macrophage",`13`="T cell",  `14`=" Proximal airway cell", `15`="Alveolar macrophage", `16`="Endothelial cell", `17`="Adipocyte")
###add cluster identity to metadata
balbc.k7m2.combined.subset$celltype.sample <- paste(Idents(balbc.k7m2.combined.subset), balbc.k7m2.combined.subset$sample, sep = "_") 
balbc.k7m2.combined.subset$celltype <- Idents(balbc.k7m2.combined.subset)

###add module scores
balbc.k7m2.combined.subset<-AddModuleScore(balbc.k7m2.combined.subset, features =Inflammation_list, name = 'Inflammation_features')
balbc.k7m2.combined.subset<-AddModuleScore(balbc.k7m2.combined.subset, features =Lung_fibrosis_list, name = 'Lung_fibrosis_features')
balbc.k7m2.combined.subset<-AddModuleScore(balbc.k7m2.combined.subset, features =TNFA_list, name = 'TNFA_features')
balbc.k7m2.combined.subset<-AddModuleScore(balbc.k7m2.combined.subset, features =ECM_degrad_list, name = 'ECM_degrad_features')
balbc.k7m2.combined.subset<-AddModuleScore(balbc.k7m2.combined.subset, features =ECM_org_list, name = 'ECM_org_features')
balbc.k7m2.combined.subset<-AddModuleScore(balbc.k7m2.combined.subset, features =Wound_response_list, name = 'Wound_response_features')
###subset epithelial cells and re-normalize
balbc.k7m2.combined.AEC=subset(balbc.k7m2.combined.subset, idents = "Distal airway cell")
balbc.k7m2.combined.AEC<- SCTransform(balbc.k7m2.combined.AEC, vars.to.regress = "percent.mt", verbose = FALSE, return.only.var.genes = FALSE)
###dim reduction
balbc.k7m2.combined.AEC<- RunPCA(balbc.k7m2.combined.AEC, verbose = FALSE)
balbc.k7m2.combined.AEC<- RunUMAP(balbc.k7m2.combined.AEC, dims = 1:30, verbose = FALSE)
###cluster
balbc.k7m2.combined.AEC<- FindNeighbors(balbc.k7m2.combined.AEC, dims = 1:30, verbose = FALSE)
balbc.k7m2.combined.AEC<- FindClusters(balbc.k7m2.combined.AEC,resolution=0.6, verbose = FALSE)
###remove contaminating non-epithelial cells-immune cells
balbc.k7m2.combined.AEC=subset(balbc.k7m2.combined.AEC, idents = c("2","3", "4", "5", "6", "7"))
###re-cluster with pure epithelial cell populations 
balbc.k7m2.combined.AEC_1<- SCTransform(balbc.k7m2.combined.AEC, vars.to.regress = "percent.mt", verbose = FALSE, return.only.var.genes = FALSE)
balbc.k7m2.combined.AEC_1<- RunPCA(balbc.k7m2.combined.AEC_1, verbose = FALSE)
balbc.k7m2.combined.AEC_1<- RunUMAP(balbc.k7m2.combined.AEC_1, dims = 1:30, verbose = FALSE)
balbc.k7m2.combined.AEC_1<- FindNeighbors(balbc.k7m2.combined.AEC_1, dims = 1:30, verbose = FALSE)
balbc.k7m2.combined.AEC_1<- FindClusters(balbc.k7m2.combined.AEC_1,resolution=0.6, verbose = FALSE)
###cluster identification based on published data sets-PMIDs 32750316,32661339,32678092 
balbc.k7m2.combined.AEC_1<-RenameIdents(balbc.k7m2.combined.AEC_1, `0`="pAEC2", `1`="DATP", `2`="AEC2", `3`="AEC2", `4`="AEC1",`5`="cAEC2")
balbc.k7m2.combined.AEC_1$celltype.sample <- paste(Idents(balbc.k7m2.combined.AEC_1), balbc.k7m2.combined.AEC_1$sample, sep = "_") 
balbc.k7m2.combined.AEC_1$celltype <- Idents(balbc.k7m2.combined.AEC_1)
###subset k7m2 macrophages
Idents(balbc.k7m2.combined.subset)="celltype.sample"
###remove "K7M2" from BALBC
balbc.k7m2.combined.subset=subset(balbc.k7m2.combined.subset, idents = "K7M2_BALBC", invert=T)
###scTransform macs 
k7m2.macs<- SCTransform(k7m2.macs, vars.to.regress = "percent.mt", verbose = FALSE, return.only.var.genes = FALSE)
###dim reduction
k7m2.macs<- RunPCA(k7m2.macs, verbose = FALSE)
k7m2.macs<- RunUMAP(k7m2.macs, dims = 1:30, verbose = FALSE)
###cluster 
k7m2.macs<- FindNeighbors(k7m2.macs, dims = 1:30, verbose = FALSE)
k7m2.macs<- FindClusters(k7m2.macs,resolution=0.4, verbose = FALSE)
k7m2.mac.markers<-FindAllMarkers(k7m2.macs, only.pos = T, logfc.threshold = 0.5, min.pct = 0.25)
###mac cluster annotation based on PMID: 35690521, 32849616, 35365629
M1up_list<-filter(all_gene_sets, gs_name=="GSE5099_CLASSICAL_M1_VS_ALTERNATIVE_M2_MACROPHAGE_UP")
M1up_list<-list(c(M1up_list$gene_symbol))
f420.macs<-AddModuleScore(f420.macs, features =M1up_list, name = 'm1_features')
phago_list<-filter(all_gene_sets, gs_name=="GOBP_POSITIVE_REGULATION_OF_PHAGOCYTOSIS")
phago_list<-list(c(phago_list$gene_symbol))
f420.macs<-AddModuleScore(f420.macs, features =phago_list, name = 'phagocytosis_features')
###remove contaminating endothelial, T cells
f420.mac.subset<-subset(f420.macs, idents = c("7", "9"), invert=T)
###mac markers based on 35690521 LA-TAM, ANGIO-TAM, IFN-TAM, INFLAM-TAM, PROLIF-TAM, Classical monocyte, nonclassical
TAM.markers<-c("Spp1", "Vegfa", "Fabp5", "Gpnmb", "Lgals3","Apoe","Pf4", "Mrc1", "Cd206","Arg1", "Cxcl1", "Cxcl2", "Cxcl3", "Il1a", "Il1b", "Inhba", "Ccl20", "Cxcl9", "Cxcl10", "Ifit1", "Ifitm1", "Tnfsf10", "Fn1", "Hmox1", "C1qa", "C1qb","Mki67","Ace", "Itgal", "Ear2", "Spn", "Itgam", "Itgax", "Cd14", "Fcgr3", "Fcgr1", "Cd68","Cd74", "Cd80", "Cd86", "Ccr2", "Cx3cr1", "Ly6c1", "Ly6c2", "H2-Ab1", "H2-DMb1")
###cluster annotation based on PMID: 35690521
dafdsfafs
