# Create SingleR References of the complete HLCA and of the core (healthy lung) HLCA
The annotated full HLCA dataset was downloaded from the Human Lung Cell Atlas project.
https://doi.org/10.1038/s41591-023-02327-2

NOTE: If you haven't already downloaded and processed, you might need to run this step on a himem node.

## Setup the environment

```{r setup}
library(tidyverse)
library(hdf5r)
library(zellkonverter)
library(scater)
library(SingleR)
# library(rrrSingleCellUtils)

# Enable parallel processing for Seurat
bp.params <- BiocParallel::MulticoreParam(workers = 30)
```

## Load and pre-process the hlca complete dataset

```{r load-hlca}
if (!file.exists("/gpfs0/home2/gdrobertslab/lab/GenRef/sc_ref_datasets/human/hlca.h5ad")) {
    # Download the annotated full dataset from the Human Lung Cell Atlas project
    # https://doi.org/10.1038/s41591-023-02327-2
    system(
        "curl https://datasets.cellxgene.cziscience.com/3ab47484-a3eb-4f6a-beea-670e1a8fc1e8.h5ad --output /gpfs0/home2/gdrobertslab/lab/GenRef/sc_ref_datasets/human/hlca.h5ad"
    )
}

if(!file.exists("LungRefs/hlca-complete.qs")) {
    # Need to invoke BPCells to handle the large object until we simplify
    system('export LD_LIBRARY_PATH="./.conda/pkgs/libstdcxx-ng-12.2.0-h46
    fd767_19/lib/libstdc++.so.6.0.30:$LD_LIBRARY_PATH"')
    hlca <- readH5AD("/gpfs0/home2/gdrobertslab/lab/GenRef/sc_ref_datasets/human/hlca.h5ad")

    # Convert Ensembl IDs to gene symbols
    conv_table <- AnnotationDbi::select(org.Hs.eg.db::org.Hs.eg.db,
        keys = rownames(hlca),
        columns = "SYMBOL",
        keytype = "ENSEMBL")
    conv_table <- conv_table[!is.na(conv_table$SYMBOL), ]
    conv_table <- conv_table[!duplicated(conv_table$SYMBOL), ]
    hlca <- hlca[conv_table$ENSEMBL, ]
    rownames(hlca) <- conv_table$SYMBOL

    qs::qsave(hlca, "LungRefs/hlca-complete.qs")

} else {
    hlca <- qs::qread("LungRefs/hlca-complete.qs")
}

    # Create an aggregated reference simplified with pseudobulk
    names(assays(hlca)) <- c("logcounts_no_soup", "logcounts")
    hlca_whole_ref <- aggregateReference(hlca,
        c(hlca$cell_type, hlca$ann_level_2),
        ncenters = 20,
        BPPARAM = bp.params,
        BSPARAM = BiocSingular::IrlbaParam())

```

# Create a healthy lung object to use in comparison to lung mets

```{r healthy-lung}
cells <- colData(hlca) %>%
    as.data.frame() %>%
    rownames_to_column(var = "cell_ids") %>%
    select(cell_ids, fresh_or_frozen, lung_condition, smoking_status) %>%
    filter(smoking_status %in% c("former", "never")) %>%
    filter(lung_condition == "Healthy") %>%
    group_by(fresh_or_frozen) %>%
    slice_sample(n = 20000) %>%
    pull(cell_ids)

healthy_lung <- hlca[, cells]




```