We are interested in how the myeloid population in the treated samples compare to that in the untreated sample. This will include looking at what subpopulations are enriched, what genes are differentially expressed, and what pathways are differentially regulated.

## Subset Out Myeloid Cells

```{r snRNA_load_merged}
if (!exists("merged_samples")) merged_samples <- qs::qread("output/rdata/nucleus/merged_samples.qs")
```

The first step to this analysis is to subset out the myeloid cells. I will do this according to the SingleR cell type predictions.

```{r subset-myeloid}
myeloid <- subset(merged_samples, idents = c(0, 12)) %>%
    process_seurat(resolution = 1)

# Visualize new umap
r_dim_plot(myeloid, split.by = "treatment")
r_dim_plot(myeloid, split.by = "treatment", group.by = "sn_cell_type")

# Look at markers
mac_genes <- c(
    "Mmp9", "Ctsk", "Nfatc1", # Osteoclast-TAMs, https://doi.org/10.1186/s41232-022-00213-x
    "Cd9", "Trem2", "Spp1", "Gpnmb", # Scar-associated, see https://doi.org/10.1038/s41586-019-1631-3
    "Ms4a7", "Selenop", # TAMs, generic
    "Ifit1", "Ifit2", "Cxcl10", # IFN
    "Il1a", "Il1b", "Fn1",  # Inflammatory
    "Marco", "Siglecf", "Itgax", # Alveolar
    "Top2a", "Hist1h1b", "Birc5",  # Cycling
    "Ly6c2", "Sell", "Ifitm6",  # cMonocyte
    "Itgal", "Ace", "Spn", # ncMonocyte
    "Cd209a", "Ccr7", "Dcstamp", # Pre-DC
    "Xcr1", "Clec9a", "Irf8") # DC1

DotPlot(myeloid,
    features = mac_genes,
    # group.by = "macs_assignment",
    cols = "RdBu",
    col.max = 1.5) +
    scale_y_discrete(limits = rev) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    coord_fixed()

myeloid2 <- RenameIdents(myeloid,
    "0" = "TAM",
    "1" = "TAM",
    "2" = "Alveolar",
    "3" = "Scar",
    "4" = "TAM",
    "5" = "TAM",
    "6" = "Cycling",
    "7" = "Pre-DC",
    "8" = "ncMonocyte",
    "9" = "Pre-DC",
    "10" = "DC1",
    "11" = "Scar",
    "12" = "IFN")

r_dim_plot(myeloid2, split.by = "treatment")

qs::qsave(myeloid, "output/rdata/nucleus/myeloid_processed.qs")
```

### Plot Markers for Each Cluster

```{r myeloid-cluster-markers}
myeloid2$macs_assignment <- Idents(myeloid2) %>%
    factor(levels = c(
        "Osteoclast",
        "Scar",
        "TAM",
        "IFN",
        "Inflammatory",
        "Cycling",
        "Alveolar",
        "Interstitial",
        "cMonocyte",
        "ncMonocyte",
        "Pre-DC",
        "DC1"))

DotPlot(myeloid2,
    features = mac_genes,
    group.by = "macs_assignment",
    cols = "RdBu",
    col.max = 1.5) +
    scale_y_discrete(limits = rev) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    coord_fixed()

```