## Looking At Whole Dataset

I'm going to see how the cell type proportions change in the whole dataset.

First I need to add james' cell types into merged_samples

I need to add the cell types from `small_epi` as well.

```{r}
merged_samples$james_cell_type <- merged_samples$coarse_cell_type
merged_samples@meta.data[colnames(myeloid),]$james_cell_type <- myeloid$james_cell_type


#now add in annotations from small_epi
merged_samples@meta.data[colnames(small_epi),]$james_cell_type <- small_epi$james_cell_type
merged_samples@meta.data <- merged_samples@meta.data %>%
    mutate(james_cell_type = recode(james_cell_type,
                                    `1` = "AEC2",
                                    `2` = "AEC1",
                                    `3` = "DATP",
                                    `4` = "primed AEC2",
                                    `5` = "FB",
                                    `6` = "endothelial",
                                    `7` = "pAEC2"))


merged_samples$james_cell_type <- ifelse(merged_samples$seurat_clusters == 2,
                                         "tumor",
                                         merged_samples$james_cell_type)

all_cts <- merged_samples@meta.data %>%
    group_by(treatment, james_cell_type) %>%
    mutate(total = n()) %>%
    select("treatment", "james_cell_type", "total") %>%
    unique() %>%
    pivot_wider(names_from = "treatment", values_from = "total") %>%
    column_to_rownames("james_cell_type")

#convert NAs to 0s
all_cts[is.na(all_cts)] <- 0

round(prop.table(as.matrix(all_cts), margin = 2), 4)

niche_cells <- subset(merged_samples, james_cell_type != "tumor") %>%
    process_seurat()

niche_cells$treatment <- recode(niche_cells$sample_id,
                            "S0291" = "Control",
                            "S0283" = "Late Nintedanib",
                            "S0284" = "Early Nintedanib")
```

Cell types that are in niche_cells now

"AEC1" = "Epithelial",
"Mes" = "Mesothelial",
"TAMs" = "Monocyte/Macrophage",
"DATP" = "Epithelial",
"AEC2" = "Epithelial",
"Mac" = "Monocyte/Macrophage",
"Endothelial" = "Endothelial",
"FB" = "Fibroblast",
"endothelial" = "Endothelial",
"Scar-TAMs" = "Monocyte/Macrophage",
"Peri" = "Smooth muscle cells",
"Cil" = "Epithelial",
"Club" = "Epithelial",
"Interstitial" = "Monocyte/Macrophage",
"T cell" = "T/NK cells",
"B cell" = "B cell",
"Epithelial" = "Epithelial",
"Mono" = "Monocyte/Macrophage",
"DC" = "Dendritic cell",
"SMC" = "Smooth muscle cells",
"pAEC2" = "Epithelial",
"primed AEC2" = "Epithelial",
"Bas" = "Granulocyte",
"Angio-TAMs" = "Monocyte/Macrophage"

## Add in These Cell Types

```{r make-final-cell-types}
niche_cells$final_cell_type <-
    recode(niche_cells$james_cell_type,
           "AEC1" = "Epithelial",
           "Mes" = "Mesothelial",
           "TAMs" = "Monocyte/Macrophage",
           "DATP" = "Epithelial",
           "AEC2" = "Epithelial",
           "Mac" = "Monocyte/Macrophage",
           "Endothelial" = "Endothelial",
           "FB" = "Fibroblast",
           "endothelial" = "Endothelial",
           "Scar-TAMs" = "Monocyte/Macrophage",
           "Peri" = "Smooth muscle cells",
           "Cil" = "Epithelial",
           "Club" = "Epithelial",
           "Interstitial" = "Monocyte/Macrophage",
           "T cell" = "T/NK cells",
           "B cell" = "B cell",
           "Epithelial" = "Epithelial",
           "Mono" = "Monocyte/Macrophage",
           "DC" = "Dendritic cell",
           "SMC" = "Smooth muscle cells",
           "cAEC2" = "Epithelial",
           "pAEC2" = "Epithelial",
           "Bas" = "Granulocyte",
           "Angio-TAMs" = "Monocyte/Macrophage")

r_dim_plot(niche_cells, group.by = "final_cell_type")
```

## Determine Proximal vs Distal Airway Cells

```{r}
epi_feats <- c("Scgb1a1", "Epcam", "Tuba1a", "Sftpd", "Ager", "Aqp5")

final_epi <- subset(niche_cells, final_cell_type == "Epithelial") %>%
    process_seurat()

r_feature_plot(final_epi, features = epi_feats, coord.fixed = TRUE)

r_dim_plot(final_epi)
```

Based on these pltos, specifically the expression of Scgb1a1 and Ager and Aqp5, I am led to believe that clusters 3 and 5 are proximal airway cells, and clusters 0, 1, 2, 4, 6, and 7 are distal airway cells.

```{r add-epi-annotations-prox-dist}
final_epi$final_cell_type <- recode(final_epi$seurat_clusters,
                                    `0` = "Distal airway cell",
                                    `1` = "Distal airway cell",
                                    `2` = "Distal airway cell",
                                    `3` = "Proximal airway cell",
                                    `4` = "Distal airway cell",
                                    `5` = "Proximal airway cell",
                                    `6` = "Distal airway cell",
                                    `7` = "Distal airway cell")

#now I need to add these back to my niche_cells object
niche_cells@meta.data[colnames(final_epi), ]$final_cell_type <-
    final_epi$final_cell_type

niche_cells$final_cell_type <- recode(niche_cells$final_cell_type,
                                      `1` = "Distal airway cell",
                                      `2` = "Proximal airway cell")

#UMAP of niche cells
r_dim_plot(niche_cells, group.by = "final_cell_type", split.by = "treatment")
```

## Determine Alveolar Macs

```{r}
mac_feats <- c("Itgam", "Fcgr1", "Cx3cr1", "Cd68", "Itgax", "Mertk")
niche_macs <- subset(niche_cells, final_cell_type == "Monocyte/Macrophage") %>%
    process_seurat()

#feature plot to differentiate alveolar macs from other macs
r_feature_plot(niche_macs, features = mac_feats, coord.fixed = TRUE, ncol = 3)

r_dim_plot(niche_macs)
```

Based on this feature plot, I'm confident that clusters 2, 3, 5, and 7 are our alveolar macrophages.

```{r add-alv-mac-annotations}
niche_macs$final_cell_type <-
    ifelse(niche_macs$seurat_clusters %in% c(2, 3, 5, 7),
           "Alveolar macrophage",
           "Monocyte/Macrophage")
```

## Add Annotations Back to niche_cells


```{r}
niche_cells@meta.data[colnames(niche_macs), ]$final_cell_type <-
    niche_macs$final_cell_type

pdf("output/figures/nucleus/niche_cell_split_umap.pdf", height = 6, width = 15)
r_dim_plot(niche_cells, group.by = "final_cell_type", split.by = "treatment")
dev.off()
```
